
# "Corsework2"
## "ufcnrtr"


### Libraries
First we are going to upload the libraries

```
library(tmap)
library(geojsonio)
library(sf)
library(raster)
library(tidyverse)
library(spData)
library(spatstat)
library(sp)
library(raster)
```

```{r include=FALSE}
library(tmap)
library(geojsonio)
library(sf)
library(raster)
library(tidyverse)
library(spData)
library(spatstat)
library(sp)
library(raster)
```



### Then we are going to read the data from Adam's Dropbox for the route.
```{r}
hunt <- geojson_read("https://www.dropbox.com/s/wa2ip35tcmt93g3/Team7.geojson?raw=1", what = "sp")
```


### Let's have a look at the data.

```{r}
tmap_mode("view")
tm_shape(hunt) +
  tm_lines (col = "green", lwd = 4)
```

### sooooo we converted sp to sf .....we measure the distance and we change the coordinate system
st_as_sf ------st means is the generic of simple features, therefore we are transforming from sp to sf

```{r}
hunt1 <- st_as_sf(hunt)
st_crs(hunt1)
```
 we should take a look at the projection
 we also can ask to the head of our data
 simple features dont work well with rasters

```{r}
head(hunt1)
```
### Let's measure the distance

```{r}
st_length(hunt1)
```
### Let's change to the national british grid

```{r}
hunt2 <- st_transform(hunt1, 27700)
st_crs
```

```
library(rgdal)
```
```{r include=FALSE}
library(rgdal)
```

### Read tube statations file as simple features

```{r}
tubestations <- st_read("https://www.doogal.co.uk/LondonStationsKML.ashx","London stations with zone information")
```
### Change to National British grid and plot, we can confirm ploting the head of the data.
### Let's plot the stations.

```{r}
stations1<-  st_transform(tubestations, 27700) 
head(stations1)
tm_shape(stations1) +
  tm_dots(col="blue")
```
###  Next thing is the 100m distance buffer for the route. 

```{r}
hunt3 <- st_buffer(hunt2, dist = 100)
class(hunt2)
tm_shape(hunt3) +
  tm_polygons(col = "red")
```
### super buffer is done
### now we should plot the tube stations 

```
library(tidyverse)
```

```{r include=FALSE}
library(tidyverse)
```
### Read the hunt addresses from a csv file and check the head of the data

```{r}
huntaddresses <-  read.csv("https://www.dropbox.com/s/2cbu2ux9ddy9c0l/huntaddresses.csv?raw=1")
head(huntaddresses)
```

### I was almost on the finish line
Harry Potter's platform is giving a headache. the platform 9 3/4 has an address from the US let's change it. 

### This is gonna work
```{r}
huntaddresses[29,3] <- 51.531427
huntaddresses[29,4] <- -0.1261330
```

### Here we transform the format of the hunt addresses to simple features.

```{r}
huntaddresses1 <-  st_as_sf(huntaddresses, coords = c("lon","lat"))
head
```
 
### Transform into a national BNG
```{r}
st_crs(huntaddresses1)
huntadresses2 <- st_set_crs(huntaddresses1, 4326)
```
### It doesnt have a projection we can use this function   st_set_crs let's have a look it has a WGS84 let's change to NBG, plot the head of the data, and plot the points.

```{r}
st_crs(huntadresses2)
huntaddresses3 <- st_transform(huntadresses2, 27700)
head(huntaddresses3)
tm_shape(huntaddresses3) +
  tm_dots(col = "green")
```

### Here comes the good the intersections.

###  So I have the hunt3 which is the buffer and huntaddresses3 which is the tube stations
### Let's do the intersection 1 for question 2

```{r}
findtubestations = st_intersection(hunt3,stations1)
```
###  Let's plot the intersection
```{r}

tm_shape(hunt3)+
  tm_polygons(col='green')+
  tm_shape(findtubestations)+
  tm_dots(col="blue")
```
###  This buffer is for question number 3

```{r}
hunt4 <- st_buffer(hunt2, dist = 300)
tm_shape(hunt4) +
  tm_polygons(col = "red")
```
###  new intersection this intersection is for the question number 3 
```{r}
findhunts = st_intersection(hunt4, huntaddresses3)
tm_shape(hunt4)+
  tm_polygons(col = "blue")+
  tm_shape(findhunts)+
  tm_dots(col="red")
```
###  clip the data to a single object
###  Let's go to the new question number 4 now we load the london wards shapefile.

```
library(rgdal)
```

```{r include=FALSE}
library(rgdal)
```

### Read the shape file with OGR, transform to simple features and change to the NBG.
```{r}
LondonWards <- readOGR("London Wards/LondonWards.shp")
LondonWards1 <- st_as_sf(LondonWards)
LondonWards2<-st_transform(LondonWards1,27700)
```

### Calculate the intersection between wards

```{r}
londonWardsIntersect<- LondonWards2[hunt2,]
qtm(LondonWards2)
qtm(londonWardsIntersect)

```



### To be more clear plot the route and the wards
```{r}
tm_shape (londonWardsIntersect)+
  tm_polygons(col = NA, alpha = 0.9)+
  tm_shape(hunt) +
  tm_lines (col = "green", lwd = 4)
``` 

### This is for question 4
### Let's do some filtering ------filter the data and get the values filter male and female expectancy
```{r}
londonWardsIntersect$MaleLE0509
londonWardsIntersect$FemaleLE05
```
### Get the maximun value
```{r}
londonwardsmax<-filter(londonWardsIntersect,MaleLE0509==max(londonWardsIntersect$MaleLE0509))
```
### Get the minimun value
```{r}
londonwardsmin<-filter(londonWardsIntersect,MaleLE0509==min(londonWardsIntersect$MaleLE0509))
print (londonwardsmax)
print (londonwardsmin)
```
### This is for question number 5
### get the mean of Life expentancy male
```{r}
londonwardsaveragemale<-group_by(londonWardsIntersect)%>%
                    summarise(mean_life = mean (MaleLE0509, na.rm = TRUE))     
print(londonwardsaveragemale)
londonwardsaveragefemale<-group_by(londonWardsIntersect)%>%
                    summarise(mean_life = mean (FemaleLE05, na.rm = TRUE))     
print(londonwardsaveragefemale)
```

### For question five let's run a K test
### Let's Transform to a Spatial Polygons


```{r}
superhunt <- huntaddresses3

```

### First we change the points from sf to sp, now we do the opposite.
```{r}
score <- sf:::as_Spatial(superhunt)
```

### Plot the the window for the K test
```{r}
window <- as.owin(LondonWards)
plot(window)
```

### Now we should create a ppp object

```{r}
huntpoints <- score
huntpoints.ppp <- ppp(x=huntpoints@coords[,1],y=huntpoints@coords[,2],window=window)
plot (huntpoints.ppp,pch=16,cex=1,main="huntpoints")
```

### Finally we run our K test
#K test
```{r}
K <- Kest(huntpoints.ppp, correction="border")
plot(K)
```



















